### 2012-10-30 3.tex
writing on 4

One crucial detail is missing from this analysis: \textsc{Backness Harmony} needs to be prevented from overwriting the [$+$\textsc{Back}] specification of disharmonic roots, either by non-derived environment blocking \citep{Mascaro1976} or structure preservation condition \citep{Kiparsky1985}. However, any condition which prevents \textsc{Backness Harmony} from overwriting underlying backness specifications will reintroduce the duplication of sequence structure and phonological generalizations; under this analysis, disharmonic roots are no longer exceptional, despite considerable evidence (reviewed below) that they are marked in Turkish.\footnote{On the other hand, it is possible to interpret the presence of a single backness specification per root as a sort of default. A precedent for this is the surface-oriented interpretation of the tonal Obligatory Contour Principle proposed by \citet[134]{Goldsmith1976} and \citet{Odden1986}, under which adjacent identical tones are automatically attributed to a single underlying tone. However, this is merely a notational variant of the rule exceptionality account in which [$+$\textsc{Backness Harmony}] is the default.} On these grounds, the underspecification analysis is rejected here in favor of the rule exceptionality account.

>>>>

One crucial detail is missing from this analysis: \textsc{Backness Harmony} needs to be prevented from overwriting the [$+$\textsc{Back}] specification of disharmonic roots, either by non-derived environment blocking \citep{Mascaro1976} or structure preservation condition \citep{Kiparsky1985}. However, any condition which prevents \textsc{Backness Harmony} from overwriting underlying backness specifications will reintroduce the duplication of sequence structure and phonological generalizations; under this analysis, disharmonic roots are no longer exceptional, despite considerable evidence (reviewed below) that they are marked in Turkish.\footnote{On the other hand, it is possible to interpret the presence of a single backness specification per root as a sort of default. A precedent for this is the surface-oriented interpretation of the tonal Obligatory Contour Principle proposed by \citet[134]{Goldsmith1976} and \citet{Odden1986}, under which adjacent identical tones are automatically attributed to a single underlying tone. However, this is merely a notational variant of the rule exceptionality account in which [$+$\textsc{Backness Harmony}] is the default.} The the underspecification analysis is rejected here on these grounds.

### 2012-10-30 4.tex
writing on 4

The baroque treatment of Latinate forms in \emph{The sound pattern of English} (\citealp{SPE}, henceforth \emph{SPE}) assumes a prefix/bound stem decomposition to simplify various morphophonological generalizations.

In a similar vein, \citet[11f.]{Aronoff1976} observes that Latinate forms which share the same bound stem also share irregular allomorphs of that stem under derivation, which is presented as evidence of prefix/bound stem decomposition.

@@

This assumption has long been made to simplify phonological \citep{SPE}, allomorphic \citep[11f.]{Aronoff1976}, and syntactic \citep{Harley2009} generalizations.

Results from lexical decision experiments also suggest that neoclassical compounds are complex. \citet{Taft1986} and \citet{Taft1975,Taft1976} find that nonce words like \emph{*re-sert}, which appear to be composed of a prefix and a bound stem, take longer to reject that non-words which lack apparent morphological structure such as *\emph{refant}). Bound stems also show frequency effects which are independent of their whole word frequency \citep{Taft1979,Taft2006}. Finally, \citet{Emmorey1989} and \citet{Forster2000} report facilitative priming between pairs like \emph{permit}-\emph{submit}, an effect suggestive of a shared morphological identity.

There is also a syntactic interaction between Latinate prefixes on verbs and the makeup of the verbal complement. \citet{Harley2009} observes that Latinate verbs do not generally participate in ditransitive, verb participle, or adjectival resultative constructions, all acceptable with similar Anglo-Saxon verbs \citep[see also][]{Gropen1989,Coppock2008}. \citeauthor{Harley2009} proposes that the Latinate prefix is introduced in a low small clause, the same position as the theme, participle, and result adjectival in the Anglo-Saxon examples.

@@

@@

There is no statistical support for any of the three static constraints proposed by \citeauthor{Pierrehumbert1994}. 
%However, it is premature to conclude that no static constraint exists. It is possible that an explicit model of ``constraint discovery'', could identify reliable static constraints,

@@

@@

\noindent
\citet{Pierrehumbert1994} does not discuss a constraint against adjacent obstruents disagreeing in voice. However, the vast majority of medial obstruents clusters in simplex wors are either uniformly voiced, as in \emph{hu}[z.b]\emph{and}, or uniformly voiceless, as in or \emph{rha}[p.s]\emph{osdy}. Hetero-voiced clusters, like those in \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos} are rare in comparison.

@@

\textsc{Nasal Place Assimilation} (\emph{SPE}:85) accounts for [n $\sim$ ŋ] allophony as well as

is responsible for the allomorphs of the \emph{im-}/\emph{in-} prefix, determined by the place of the following consonant (though there are additional complexities; see \citealt{Borowsky1986}:?, \citealt{Halle1985a}, \citealt{Jensen2000}).

@@

@@

@@

\paragraph{Degemination}

The final alternation found in English medial clusters is the simplification of geminates characteristic of ``level I'' morphology. Verb roots ending in /d/ which select the irregular /-t/ past tense suffix exhibit degemination (e.g., \emph{bend}/\emph{ben}[t], \emph{build}/\emph{buil}[t]).

A final alternation targeting English syllable contact clusters is the simplification of geminates certain affixes.

The deadjectival suffix \emph{-ly} does not result in a geminate \emph{l} when it attaches to

krarely produces a geminate when it attaches to /l/-final roots (e.g., \emph{norma}[l]\emph{y}, cf. \emph{calm}[l]\emph{y}).

More complicated cases of \textsc{Degemination} can be found in Latinate prefix allomorphy (\emph{SPE}:148; \citealp[102]{Borowsky1986}).

@@

@@

>>>>

The baroque treatment of Latinate forms in \emph{The sound pattern of English} (\citealp{SPE}, henceforth \emph{SPE}) assumes a prefix/bound stem decomposition to simplify various morphophonological generalizations. In a similar vein, \citet[11f.]{Aronoff1976} observes that Latinate forms which share the same bound stem also share irregular allomorphs of that stem under derivation.

@@

\noindent
\citeauthor{Aronoff1976} takes this to be evidence that \emph{adhere} and \emph{cohere}, for instance, share a bound stem.

There is also a syntactic interaction between Latinate prefixes on verbs and the makeup of the verbal complement. \citet{Harley2009} observes that Latinate verbs do not generally participate in ditransitive, verb participle, or adjectival resultative constructions, all acceptable with similar Anglo-Saxon verbs \citep[see also][]{Gropen1989,Coppock2008}. 
%\citeauthor{Harley2009} proposes that the Latinate prefix is introduced in a low small clause, the same position as the theme, participle, and result adjectival in the Anglo-Saxon examples.

@@

Results from lexical decision experiments also suggest that neoclassical compounds are complex. \citet{Taft1975,Taft1976} and \citet{Taft1986} find that nonce words like \emph{*re-sert}, which appear to be composed of a prefix and a bound stem, take longer to reject that non-words which lack apparent morphological structure such as *\emph{refant}). Bound stems also show frequency effects which are independent of their whole word frequency \citep{Taft1979,Taft2006}. Finally, \citet{Emmorey1989} and \citet{Forster2000} report facilitative priming between pairs like \emph{permit}-\emph{submit}, an effect thought to indicate of morphological relatedness.

@@

So as to account for the 81\% of ``possible'' but unattested clusters, \citet{Pierrehumbert1994} proposes three constraints on English medial clusters. These constraints are ``static'' in the sense that they have no analogues in the morphophonemics of English. The investigation considers clusters of all sizes, not merely the triconsontal clusters studied by \citeauthor{Pierrehumbert1994}, as she gives no justification for excluding biconsontal or quadraconsonantal clusters.

\paragraph{Dorsal-labial clusters} \citet[173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars'', noting that velar-velar clusters are independently excluded by constraints on geminates (see \S?). Two-consonant velar-labial clusters are found in words like \emph{a}[k.m]\emph{e}, \emph{ru}[ɡ.b]\emph{y}, or \emph{pi}[ɡ.m]\emph{ent} occur. Table \ref{dltab} shows that velar-labial clusters are somewhat less likely to occur than  are velar-coronal clusters (e.g., \emph{ve}[k.t]\emph{or}), but such underrepresentation is not unlikely to occur by chance.

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 25       & 91         & 22\%       & \multirow{2}{*}{0.106} \\
violating  &  4       & 40         &  9\%       \\
\bottomrule
\end{tabular}
\caption{FIXME}
\label{dltab}
\end{table}

\paragraph{Coronal codas} \citet[175]{Pierrehumbert1994} claims that ``clusters with a coronal obstruent in the coda do not occur'', while noting exceptions like \emph{a}[nt.l]\emph{er}, \emph{ke}[s.tr]\emph{el} and \emph{oi}[nt.m]\emph{ent}. Table \ref{cctab} shows that coda coronal obstruent clusters are not significantly less likely to occur than non-coronal obstruent clusters (e.g., \emph{re}[p.t]\emph{ile}). Restricting attention to triconsonantal clusters does not produce a significant lexical effect, either ($p = 0.129$).

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 56       & 304        & 15\%       & \multirow{2}{*}{0.430} \\
violating  & 37       & 243        & 13\%       \\
\bottomrule
\end{tabular}
\caption{FIXME}
\label{cctab}
\end{table}

\noindent

\paragraph{ABA clusters} \citet[][176]{Pierrehumbert1994} notes the ``lack of clusters with identical first and third elements'', ignoring presence or absence of voicing. Despite the fact that the corpus contains no exceptions to this generalization, these \textsc{ABA} clusters are not significantly less common than other triconsonantal and quadraconsonantal clusters (Table \ref{abatab}).

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 47       & 512        &  8\%       & \multirow{2}{*}{0.250} \\
violating  &  0       &  25        &  0\%       \\
\bottomrule
\end{tabular}
\caption{FIXME}
\label{abatab}
\end{table}

\paragraph{Summary} There is no statistical support for \citeauthor{Pierrehumbert1994}'s static constraints.

@@

FIXME

@@

\noindent
\citet{Pierrehumbert1994} does not discuss a constraint against adjacent obstruents disagreeing in voice. As shown in Table \ref{ovatab} however, the vast majority of medial obstruents clusters in simplex words are either uniformly voiced, as in \emph{hu}[z.b]\emph{and}, or uniformly voiceless, as in or \emph{rha}[p.s]\emph{osdy}, and hetero-voiced clusters, like those in \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos} are rarer than would be expected from chance.

@@

\textsc{Nasal Place Assimilation} (\emph{SPE}:85) accounts for [n $\sim$ ŋ] allophony and Latinate prefix allomorphy, such as that of the ``negative'' refix \emph{im-}/\emph{in-}.\footnote{\citealt{Borowsky1986}:FIXME and \citealt{Jensen2000} propose additional restrictions on \textsc{Nasal Place Assimilation}.}

@@

%Furthermore, \citet[228]{Myers1993} reports speech errors which create nasal-obstruent clusters and undergo \textsc{Nasal Place Assimilation}; e.g., \emph{ra}[nd] \emph{orker}, intended \emph{ra}[ŋk] \emph{order}.

\noindent
The rule is formalized below.

@@

As shown in Table \ref{npatab}, virtually all (94\%) clusters of a nasal coda followed by a homorganic obstruent (e.g., \emph{pi}[m.p]\emph{le}, \emph{sta}[n.z]\emph{a}, \emph{mo}[ŋ.k]\emph{ey}) are attested. As \citet[175]{Pierrehumbert1994} observes, heterorganic clusters, like those \emph{pli}[m.s]\emph{oll} or \emph{scri}[m.ʃ]\emph{aw}, do occur, but they are significantly more rare.

@@

\paragraph{Degemination} The final alternation found in English medial clusters is the simplification of geminates which is characteristic of ``level I'' morphology, and is found in the irregular /-t/ past tense (e.g., \emph{bend}/\emph{ben}[t], \emph{build}/\emph{buil}[t]) and in \emph{-ly} deadjectival derivatives (e.g., \emph{norma}[l]\emph{y}, cf. \emph{calm}[l]\emph{y}), and Latinate prefix allomorphy (\emph{SPE}:148; \citealp[102]{Borowsky1986}). \textsc{Degemination} is formalized here as a rule deleting the first of two identical segments, ignoring voice.

@@

\noindent
As can be seen from Table \ref{degemtab}, the sample includes no sequences of identical segments, or of identical segments differing only in voice, something highly unlikely to arise by chance.

@@

\paragraph{Summary} All three of \emph{SPE} rules targeting medial clusters have a robust effect in constraining the inventory of possible word-medial syllable clusters.
